Reasoning about the rigor of perl programs: \\ \large gaps, limits to understanding, and means to ameliorate

### 2024-05-21 sciperl/paper.tex
Yahoo, citations are working

\documentclass[twoside]{article}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage[dvipsnames]{xcolor}
\hypersetup{
     colorlinks=true,
     allcolors=NavyBlue
}
\usepackage{listings}
\lstset{language=Perl}

@@

\section{INTRODUCTION}

@@

1. rigorous testing of any individual *instruction* within a stack over it's expected domain of inputs, and a sampling of data outside of that is valuable.
2. The interaction of an instruction with its immediate follower and predecessor must also be well understood; these multi-instruction sections can indeed be treated as though they are themselves more complicated instructions.

@@

Building such a tree is how PPI, and the tools which rely on it such as Perl::Critic work.
Alternatively you can inspect the "intermediate form" which the interpreter itself builds via the B::Deparse, or similar techniques as Devel::Cover and Devel::NYTProf leverage.
However just because you *can* reason about perl programs in the same fashion as any other programming language, does not guarantee avenues of inquiry are exhaustively explored.

@@

It's usually fairly straightforward to get 100% statement coverage on greenfield code.
However this rarely translates into software quality, as the test coverage on any of the deps in the stack remains opaque.
How do we identify if the dependencies we use are well covered *where we use them*?
If we can, the same techniques can be cross-referenced with the relative importance of the module vis the "CPAN River" to identify where the most high-impact test coverage improvements could be made in the perl ecosystem.

@@

While we could always wrap CPAN, ideally a CPAN::Plugin either pre or post test would be ideal.
Nevertheless, given the wide variety of means that CPAN Modules might be installed, any solution here would have to build a coverage index on-demand, and cache it for later use, likely as a plugin to Devel::Cover instead.
cpancover *appears* to do this, but is currently broken and appears unmaintained.

@@

Since the primary programming paradigm is *not* "Klingon Programming" where it is better to die() than return() in failure, there is no guarantee that environmental problems are caught.
This is why a great many Perl::Critic rules about using eval {...;1;} or do..., `my $pid=fork(); die unless defined $pid` and so on exist.
Some means of actual fault injection should also be taken into account with regard to coverage.

@@

This is actually a quite complicated subject.  As the founder of Anththesis (https://antithesis.com/blog/is_something_bugging_you/) puts it:

the $100 bill is surrounded by an entire field of poisonous spike-pits, nay, a continent of them.

Aside from building something which can inject the *astonishingly vast array* of such faults, there is still the problem of integrating this into perl's testing harness, and reporting coverage to do.

@@

\section{Conclusion}

@@

>>>>

\documentclass[twoside]{article}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{csquotes}
\usepackage[T1]{fontenc}
\usepackage[dvipsnames]{xcolor}
\hypersetup{
     colorlinks=true,
     allcolors=NavyBlue
}
\usepackage{listings}
\lstset{language=Perl}

\begin{filecontents}{paper.bib}
    @misc{antithesis,
        title   = "Is something Bugging you?",
        author  = "Will Wilson",
        url     = {https://antithesis.com/blog/is_something_bugging_you/},
        year    = 2024
    }
\end{filecontents}

@@

\newpage
\section{INTRODUCTION}

@@

1. rigorous testing of any individual \textit{instruction} within a stack over it's expected domain of inputs, and a sampling of data outside of that is valuable.
2. The interaction of an instruction with its immediate follower and predecessor must also be well understood; these multi-instruction sections can indeed be treated as though they are themselves more complicated instructions.

@@

Building such a tree is how PPI, and the tools which rely on it such as Perl::Critic work.
Alternatively you can inspect the "intermediate form" which the interpreter itself builds via the B::Deparse, or similar techniques as Devel::Cover and Devel::NYTProf leverage.
However just because you \textit{can} reason about perl programs in the same fashion as any other programming language, does not guarantee avenues of inquiry are exhaustively explored.

@@

It's usually fairly straightforward to get 100% statement coverage on greenfield code.
However this rarely translates into software quality, as the test coverage on any of the deps in the stack remains opaque.
How do we identify if the dependencies we use are well covered \textit{where we use them}?
If we can, the same techniques can be cross-referenced with the relative importance of the module vis. the "CPAN River" to identify where the most high-impact test coverage improvements could be made in the perl ecosystem.

@@

While we could always wrap CPAN, ideally a CPAN::Plugin either pre or post test would be ideal.
Nevertheless, given the wide variety of means that CPAN Modules might be installed, any solution here would have to build a coverage index on-demand, and cache it for later use, likely as a plugin to Devel::Cover instead.
cpancover \textit{appears} to do this, but is currently broken and appears unmaintained.

@@

Since the primary programming paradigm is \textit{not} "Klingon Programming" (where it is better to \textit{die()} than \textit{return()} in failure), there is no guarantee that environmental problems are caught.
This is why a great many Perl::Critic rules about using eval {...;1;} or do..., `my $pid=fork(); die unless defined $pid` and so on exist.
Some means of actual fault injection should also be taken into account with regard to coverage.

@@

This is actually a quite complicated subject.  As the founder of Antithesis puts it \cite{antithesis}:
\begin{displayquote}
the \$100 bill is surrounded by an entire field of poisonous spike-pits, nay, a continent of them.
\end{displayquote}
Aside from building something which can inject the \textit{astonishingly vast array} of such faults, there is still the problem of integrating this into perl's testing harness, and reporting coverage to do.

@@

\newpage
\section{Conclusion}

@@

\newpage
\bibliographystyle{vancouver}
\bibliography{paper}
